---
title: "Exercise 1 Statistical Bioinformatics"
format: html
editor: visual
---

# Exercise 1

The goal of this exercise is to sample a log-normal distribution and to plot this distribution with histograms and a violin plot. An expected value of `mu = 5` and a standard deviation of `sigma = 0.5` are chosen.

```{r}
# imports
library(ggplot2)
```

## Sample log-normal distribution

```{r}
# Sampling 100 values
set.seed(12)

mu = 5
sigma = 0.5

sample_values <- rlnorm(100, meanlog = mu, sdlog = sigma)
head(sample_values)
```

## Histogram

First a histogram with normal scales is plotted.

```{r}
data <- data.frame(sample_values)
hist_linear <- ggplot(data, aes(x=sample_values)) + 
  geom_histogram(binwidth=5, fill="lightblue", color="white", alpha=0.7) +
  labs(title="Histogram of Sampled Values", 
       x="Value", 
       y="Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

print(hist_linear)
```

# Histogram with log scale

## Since the data is sampled from a log-normal distribution, it makes sense to plot a logarithmic x-axis in the histogram. Now expectation value and standard deviation are visible.

```{r}
log_sample_values <- log(sample_values)
data_log = data.frame(log_sample_values)

hist_log <- ggplot(data_log, aes(x=log_sample_values)) + 
  geom_histogram(binwidth=0.1, fill="blue", color="white", alpha=0.7) +
  labs(title="Histogram of Sampled Values with Log-scaled X-axis", 
       x="Value (Log scale)", 
       y="Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

print(hist_log)
```

# Violin plot

The distribution of the data can also be shown with a violin plot.

```{r}
violin_pl <- ggplot(data, aes(x = factor(1), y = sample_values)) +
  geom_violin(fill = "darkblue", color = "white", alpha = 0.7) +
  labs(title = "Violin Plot of Sampled Values", 
       x = "Data", 
       y = "Value") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

print(violin_pl)
```
